---
title: "Digital Health Exam Markdown"
author: "Chioma Sandra Achugonye"
date: "2024-02-06"
output: 
  rmdformats::readthedown
---



## Installing packages

```{r}
#install.packages("rmdformats")
#install.packages("skimr")
#install.packages("naniar")
#install.packages("ggplot2")
#install.packages("dplyr")
#install.packages("tidyr")
#install.packages("plotly")
#install.packages("forecats")
#install.packages("corrplot")
#install.packages("tibble")
```

```{r}
library(rmdformats)
library(tibble)
library(corrplot)
library(plotly)
library(forcats)
library(skimr)
library(naniar)
library(ggplot2)
library(dplyr)
library(tidyr)
```




## Data Importation


## Setting working directory

```{r}
setwd("C:/Users/User/Desktop/Global health R/data_ICU/Digital health assignment")
```

## Importing the Intensive Care Unit (ICU) bed data from US 

```{r}
ICU <- read.csv("ICU_Bed.csv", header = TRUE)
```




## Data exploration (Primary)


## Data overview

```{r}
head(ICU)
summary(ICU)
```

## Display a vertical preview of the dataset and dimension of the dataframe

```{r}
dim(ICU)
glimpse(ICU)
```

## Determine the data types of a data frame's columns

```{r}
sapply(ICU, class)
```

## Count distinct Values in All Columns

```{r}
sapply(ICU, function(x) n_distinct(x))
```




## Data exploration (Secondary)


## Displays most of the numerical attributes from summary

```{r}
skim(ICU)
gg_miss_fct(x=ICU, fct=State)
is.null(ICU)
```

## Unique state and county names 

```{r}
length(unique(ICU$State))
length(unique(ICU$County))
```




## Data Analysis (Correlation)


## Correlation matrix

```{r, echo=FALSE}
ICU_bed <- select(ICU,ICU.Beds, Total.Population,Population.Aged.60.,Percent.of.Population.Aged.60.)
cor_matrix <- cor(ICU_bed)
corrplot(cor_matrix, method = 'number', order = 'alphabet')
```




## Data Analysis (State level)


## Number of counties per state

```{r, echo=FALSE}
no_of_counties <- ICU %>% count(State) %>% arrange(n)
no_of_counties
no_of_counties %>% 
  select(State, n) %>% 
  mutate(State = fct_reorder(State, n)) %>% 
  ggplot(aes(State, n, fill=State)) +
  geom_text(aes(label=n), vjust=0.2, hjust=-0.15) +
  geom_col()+
  coord_flip()
```

## Number of ICU beds per state

```{r, echo=FALSE}
no_of_ICU_Bed <- ICU %>% group_by(State) %>% 
  summarise(ICU.Beds = sum(ICU.Beds))
no_of_ICU_Bed
no_of_ICU_Bed %>% 
  select(State, ICU.Beds) %>% 
  mutate(State = fct_reorder(State, ICU.Beds)) %>% 
  ggplot(aes(State, ICU.Beds, fill=State)) +
  geom_text(aes(label=ICU.Beds), vjust=0.2, hjust=-0.15) +
  geom_col()+
  coord_flip()
```

## Total population per state

```{r, echo=FALSE}
Total_population <- ICU %>% group_by(State) %>% 
  summarise(Total.Population = sum(Total.Population))
Total_population
Total_population %>% 
  select(State, Total.Population) %>% 
  mutate(State = fct_reorder(State, Total.Population)) %>% 
  ggplot(aes(State, Total.Population, fill=State)) +
  geom_text(aes(label=Total.Population), vjust=0.2, hjust=-0.15) +
  geom_col()+
  coord_flip()
```

## Total population aged 60 and above per state

```{r, echo=FALSE}
Total_population_60 <- ICU %>% group_by(State) %>% 
  summarise(Population.Aged.60. = sum(Population.Aged.60.))
Total_population_60
Total_population_60 %>% 
  select(State, Population.Aged.60.) %>% 
  mutate(State = fct_reorder(State, Population.Aged.60.)) %>% 
  ggplot(aes(State, Population.Aged.60., fill=State)) +
  geom_text(aes(label=Population.Aged.60.), vjust=0.2, hjust=-0.15) +
  geom_col()+
  coord_flip()
```

## Total population aged 60 per each ICU bed and above per state

```{r, echo=FALSE}
Total_population_60_per_bed <- ICU %>% group_by(State) %>% 
  summarise(Residents.Aged.60..Per.Each.ICU.Bed = sum(Residents.Aged.60..Per.Each.ICU.Bed))
Total_population_60_per_bed
Total_population_60_per_bed %>% 
  select(State, Residents.Aged.60..Per.Each.ICU.Bed) %>% 
  mutate(State = fct_reorder(State, Residents.Aged.60..Per.Each.ICU.Bed)) %>% 
  ggplot(aes(State, Residents.Aged.60..Per.Each.ICU.Bed, fill=State)) +
  geom_text(aes(label=Residents.Aged.60..Per.Each.ICU.Bed), vjust=0.2, hjust=-0.15) +
  geom_col()+
  coord_flip()
```




## Data Analysis (County level)


## Total number of counties

```{r}
Total_no_of_counties <- length(unique(ICU$County))
Total_no_of_counties
```

## Counties with highest and lowest number of ICU beds

```{r}
highest_ICU_beds <- ICU[order(ICU$ICU.Beds, decreasing = TRUE), ]
highest_ICU_beds_5 <- highest_ICU_beds[1:5,]
highest_ICU_beds_5

lowest_ICU_beds <- ICU[order(ICU$ICU.Beds, decreasing = FALSE), ]
lowest_ICU_beds_5 <- lowest_ICU_beds[1:5,]
lowest_ICU_beds_5
```

## Counties with highest and lowest total population

```{r}
highest_population <- ICU[order(ICU$Total.Population, decreasing = TRUE), ]
highest_population <- highest_population[1:5,]
highest_population

lowest_population <- ICU[order(ICU$Total.Population, decreasing = FALSE), ]
lowest_population <- lowest_population[1:5,]
lowest_population
```

## Counties with highest and lowest population aged 60 and above

```{r}
highest_population_60 <- ICU[order(ICU$Population.Aged.60., decreasing = TRUE), ]
highest_population_60 <- highest_population_60[1:5,]
highest_population_60

lowest_population_60 <- ICU[order(ICU$Population.Aged.60., decreasing = FALSE), ]
lowest_population_60 <- lowest_population_60[1:5,]
lowest_population_60
```

## Counties with highest and lowest percent population aged 60 and above relative to total population

```{r}
percentage_highest_population_above_60 <- ICU[order(ICU$Percent.of.Population.Aged.60., decreasing = TRUE), ]
percentage_highest_population_above_60 <- percentage_highest_population_above_60[1:5,]
percentage_highest_population_above_60

percentage_lowest_population_above_60 <- ICU[order(ICU$Percent.of.Population.Aged.60., decreasing = FALSE), ]
percentage_lowest_population_above_60 <- percentage_lowest_population_above_60[1:5,]
percentage_lowest_population_above_60
```

## Counties with highest and lowest ratio of population aged 60 compared to total population

```{r}
highest_ratio_population_above_60 <- ICU[order(ICU$Residents.Aged.60..Per.Each.ICU.Bed, decreasing = TRUE), ]
highest_ratio_population_above_60 <- highest_ratio_population_above_60[1:5,]
highest_ratio_population_above_60

lowest_ratio_population_above_60 <- ICU[order(ICU$Residents.Aged.60..Per.Each.ICU.Bed, decreasing = FALSE), ]
lowest_ratio_population_above_60 <- lowest_ratio_population_above_60[1:5,]
lowest_ratio_population_above_60
```

